Quantum Interval-Valued Probability: Contextuality and the Born Rule

### 2017-09-22 prop_letter.tex
prop_letter.tex \section{Gleason} Updating...

While the finite-precision extension of Kochen-Specker theorem asserts
the non-existence of certain kind of QIVPMs, we also want to understand
what a QIVPM could be when it exists. A conventional quantum probability
measure~$\mu$ can be easily be constructed using the Born rule when
one knows the current state of the system~\citep{Born1983,peres1995quantum,544199,Jaeger2007},
and Gleason's theorem asserts that the only state consistent with
$\mu$ on all projectors is the one used to construct $\mu$~\citep{gleason1957,Redhead1987-REDINA,peres1995quantum}.
Although the Born rule could incorporate the idea of finite precision
easily, there is a QIVPM such that every state can only consistent
with it on some of projectors. However, if we only focus on the QIVPM~$\bar{\mu}$
induced by a state~$\rho$ according to the Born rule, called the
Born QIVPMs, we found the states consistent with $\bar{\mu}$ will
be closer and closer to $\rho$ while shrinking the length of intervals.

The first step toward a finite precision Born rule is to find a way
replacing the computed probability by an interval. This can be done
by carefully picking a set of intervals and a rule to select the correct
interval when intervals are overlapping. For example, consider the
following collection of intervals: 
\begin{equation}
\begin{aligned}\mathscr{I}_{n}= & \set{\left[\tfrac{k-1}{2n},\tfrac{k+1}{2n}\right]\cap\left[0,1\right]}{k=0,\ldots,2n}\\
 & \cup\{\imposs,\necess\}
\end{aligned}
\end{equation}
which are overlapping intervals with length~$1/n$ except the intervals
with length~$\left(2n\right)^{-1}$ at each end. To select the correct
interval from these overlapping intervals~$[\ell,r]$, we define
the non-overlapping interior part of a interval, $\textrm{Int}\left([\ell,r]\right)$,
as in Table~\ref{tab:I2}.
\begin{table}
\caption{\label{tab:I2}Intervals in $\mathscr{I}_{n}$ and their interior
with an concrete example for $\mathscr{I}_{2}$. Although intervals~$[\ell,r]$
are overlapping, their interior~$\textrm{Int}\left([\ell,r]\right)$
are non-overlapping.}

@@

\yutsung{Not sure what is the best way embedding the condition for
$k$ in the table...}
\end{table}
Then, given a real-valued probability~$x$, we will select a interval~$\iota_{n}\left(x\right)=[\ell,r]$
if $x\in\textrm{Int}\left([\ell,r]\right)$.

After we have the rule to replace a real number by an interval, we
can induce a QIVPM from a state according to the following theorem.

@@

\noindent Given $x_{0}$, $x_{1}$, and $x_{0}+x_{1}\in\left[0,1\right]$,
it is easy to verify $\iota_{n}\left(x_{0}+x_{1}\right)\subseteq\iota_{n}\left(x_{0}\right)+\iota_{n}\left(x_{1}\right)$.
This implies Eq.~(\ref{eq:QuantumInterval-valuedProbability-Inclusion})
when $P_{0}$ and $P_{1}$ are orthogonal. When $P_{0}$ and $P_{1}$
are not orthogonal on a three-dimensional Hilbert space, either one
of them projects onto the subspace which the other one project onto,
or their complements are orthogonal which reduced the case we discussed~\citep{HOSTunpublished}.
Therefore, $\iota_{n}\circ\mu_{\rho}^{\mathrm{B}}$ is indeed a QIVPM.
For example, when $n=2$, two states~$\mathbb{1}/3$ and $\rho_{0}$
can induce two QIVPMs~$\iota_{2}\circ\mu_{\mathbb{1}/3}^{\mathrm{B}}$
and $\iota_{2}\circ\mu_{\rho_{0}}^{\mathrm{B}}$, where $\mathbb{1}/3=\left(\proj{0}+\proj{1}+\proj{2}\right)/3$
and $\rho_{0}=\frac{7}{24}\cdot\proj{0}+\frac{1}{3}\cdot\proj{1}+\frac{3}{8}\cdot\proj{2}$.
The interval-valued probability of the induced QIVPMs, $\iota_{2}\left(\mu_{\mathbb{1}/3}^{\mathrm{B}}\left(P\right)\right)$
and $\iota_{2}\left(\mu_{\rho_{0}}^{\mathrm{B}}\left(P\right)\right)$,
for any projector $P$ in the minimal subspace of events containing
$\proj{0}$, $\proj{1}$, and $\proj{2}$, $\events_{C}$, are computed
straightforwardly in Table~\ref{tab:Compute-the-QIVPM}.
\begin{table}
\caption{\label{tab:Compute-the-QIVPM}Compute $\iota_{2}\circ\mu_{\mathbb{1}/3}^{\mathrm{B}}$
and $\iota_{2}\circ\mu_{\rho_{0}}^{\mathrm{B}}$ based on $\mu_{\mathbb{1}/3}^{\mathrm{B}}$
and $\mu_{\rho_{0}}^{\mathrm{B}}$ on $\events_{C}$.}

>>>>

While the finite-precision extension of the Kochen-Specker theorem
asserts the non-existence of a certain kind of QIVPMs, we also want
to understand what a QIVPM could be when it exists. A conventional
quantum probability measure~$\mu$ can easily be constructed using
the Born rule when one knows the current state of the system~\citep{Born1983,peres1995quantum,544199,Jaeger2007}.
Gleason's theorem asserts that the only state consistent with $\mu$
on all projectors is the one used to construct $\mu$~\citep{gleason1957,Redhead1987-REDINA,peres1995quantum}.
Although the Born rule could incorporate the idea of finite precision
easily, there is a QIVPM such that every state can only consistent
with it on some of the projectors. However, if we only focus on the
Born QIVPMs~$\bar{\mu}$, that is, QIVPM's induced by a state~$\rho$
according to the Born rule, we find that if we monotonically shrink
the length of intervals, the states consistent with $\bar{\mu}$ will
be closer and closer to $\rho$, and vise visa.

The first step toward a finite precision Born rule is to find a way
of replacing the computed probability by an interval. This can be
done by carefully picking a set of intervals and a rule to select
the correct interval when intervals are overlapping. For example,
consider the following collection of intervals: 
\begin{equation}
\begin{aligned}\mathscr{I}_{n}= & \set{\left[\tfrac{k-1}{2n},\tfrac{k+1}{2n}\right]\cap\left[0,1\right]}{k=0,\ldots,2n}\\
 & \cup\{\imposs,\necess\}
\end{aligned}
\end{equation}
which are overlapping intervals with length~$1/n$ except for the
intervals with length~$\left(2n\right)^{-1}$ at each end. To select
the correct interval from these overlapping intervals~$[\ell,r]$,
we define the non-overlapping interior part of a interval, $\textrm{Int}\left([\ell,r]\right)$,
as in Table~\ref{tab:I2}.
\begin{table}
\caption{\label{tab:I2}Intervals in $\mathscr{I}_{n}$ and their interior
with an concrete example for $\mathscr{I}_{2}$. Although intervals~$[\ell,r]$
are overlapping, their interior~$\textrm{Int}\left([\ell,r]\right)$
are non-overlapping.}

@@

\yutsung{Not sure what is the best way embedding the condition for
$k$ in the table...}
\end{table}
Then, given a real-valued probability~$x$, we will select an interval~$\iota_{n}\left(x\right)=[\ell,r]$
if $x\in\textrm{Int}\left([\ell,r]\right)$.

After we have the rule to replace a real number by an interval, we
can induce a QIVPM from a state according to the following theorem:

@@

\noindent Given $x_{0}$, $x_{1}$, and $x_{0}+x_{1}\in\left[0,1\right]$,
it is easy to verify that
\begin{equation}
\iota_{n}\left(x_{0}+x_{1}\right)\subseteq\iota_{n}\left(x_{0}\right)+\iota_{n}\left(x_{1}\right)\,.
\end{equation}
This implies Eq.~(\ref{eq:QuantumInterval-valuedProbability-Inclusion})
when $P_{0}$ and $P_{1}$ are orthogonal. When $P_{0}$ and $P_{1}$
are not orthogonal on a three-dimensional Hilbert space, then we have
two possibilities:
\begin{itemize}
\item Their complements are orthogonal reducing to the case we discussed.
\item One of them projects onto the subspace which the other one projects
onto.
\end{itemize}
\noindent Therefore, $\iota_{n}\circ\mu_{\rho}^{\mathrm{B}}$ is indeed
a QIVPM. For example, when $n=2$, two states~$\mathbb{1}/3$ and
$\rho_{0}$ can induce two QIVPMs~$\iota_{2}\circ\mu_{\mathbb{1}/3}^{\mathrm{B}}$
and $\iota_{2}\circ\mu_{\rho_{0}}^{\mathrm{B}}$, where $\mathbb{1}/3=\left(\proj{0}+\proj{1}+\proj{2}\right)/3$
and $\rho_{0}=\frac{7}{24}\cdot\proj{0}+\frac{1}{3}\cdot\proj{1}+\frac{3}{8}\cdot\proj{2}$.
The interval-valued probability of the induced QIVPMs, $\iota_{2}\left(\mu_{\mathbb{1}/3}^{\mathrm{B}}\left(P\right)\right)$
and $\iota_{2}\left(\mu_{\rho_{0}}^{\mathrm{B}}\left(P\right)\right)$,
for any projector $P$ in the minimal subspace of events containing
$\proj{0}$, $\proj{1}$, and $\proj{2}$, $\events_{C}$, are computed
straightforwardly in Table~\ref{tab:Compute-the-QIVPM}.
\begin{table}
\caption{\label{tab:Compute-the-QIVPM}Compute $\iota_{2}\circ\mu_{\mathbb{1}/3}^{\mathrm{B}}$
and $\iota_{2}\circ\mu_{\rho_{0}}^{\mathrm{B}}$ based on $\mu_{\mathbb{1}/3}^{\mathrm{B}}$
and $\mu_{\rho_{0}}^{\mathrm{B}}$ on $\events_{C}$.}
